A Plasma Torus Around a Young Low-Mass Star

### 2025-03-19 paper/ms.tex
update

\begin{figure}[!t]
  \centering
  \includegraphics[width=\textwidth]{figures/sf5.pdf}
  \caption{Time evolution of select regions in the Keck/HIRES
  spectra from 17 February 2025.
  The horizontal axis shows $\Delta v \equiv c (\lambda -
  \lambda_0)/\lambda_0$, the relative velocity centered on each line's
  rest wavelength, normalized by the stellar equatorial velocity.
  }
  \label{fig:hirescuts}
\end{figure}

@@

Finally, the Li\,\textsc{I} 6708\,\AA\ doublet, which shows no obvious
absorption, as well as the broad K\,\textsc{I} 7699\,\AA\ feature are
both shown in Figure~\ref{fig:hirescuts}.  In the latter, narrow
telluric absorption features overlap the blue wing of the line.
Neither of these regions shows any notable variability.

@@

Circumstellar H$\alpha$ emission might be sourced either from
scattering of stellar H$\alpha$ photons, or from recombination.  We
neglect scattering because we see the amplitude of the circumstellar
emission to vary by a factor of $\approx$5, which is not mirrored in
the chromospheric line core.  The volume emissivity under case B
recombination (in which Lyman-series photons are trapped by high
optical depth, so that recombinations cascade through H$\alpha$ and
other Balmer lines) can be written as
\begin{equation}
  j_{\rm H\alpha} = n_{\rm e} n_{\rm p} \alpha^{\rm eff}_{\rm H\alpha}(T) h \nu_{\rm H\alpha},
\end{equation}
where $n_{\rm e}$ and $n_{\rm p}$ are the electron and proton densities,
and $\alpha^{\rm eff}_{\rm H\alpha}(T)$ is the effective recombination
coefficient, i.e.\ the number of recombinations per unit volume, per
unit time.
For hydrogen with temperatures between 1,000 and 10,000\,K with
densities $n_{\rm H}$$\sim$10$^{10}$\,cm$^{-3}$,
$\alpha^{\rm eff}_{\rm H\alpha}$ is typically on the order of $10^{-12}$ to
$10^{-13}$\,cm$^3$\,s$^{-1}$ \cite{Hummer1987}.
Neglecting the effects of atoms other than hydrogen, we can assume
an ionization fraction $x$, such that $n_{\rm e} = n_{\rm p} = x
n_{\rm H}$, for $n_{\rm H}$ the hydrogen number density.  Let $L_{\rm
H\alpha} = j_{\rm H\alpha} V$, for $V$ the volume of the emitting
hydrogen.
The luminosity of circumstellar hydrogen emission, $L_{\rm H\alpha}$,
is an observable: our SED fitting routine yields
$L_\star$$\approx$0.012\,$L_\odot$, which gives
$L_{\rm H\alpha},\star$$\approx$1.0$\times$$10^{28}$erg\,s$^{-1}$, and
the luminosity of the clump $L_{\rm H\alpha}$ is of order one tenth
that of the star.
Assuming an emitting sphere of radius $r$, we can write
\begin{equation}
  n_H = \frac{1}{x}
  \left( 
    \frac{L_{\rm H\alpha}}{ \alpha^{\rm eff}_{\rm H\alpha}(T) h \nu_{\rm H\alpha} V }
  \right)^{1/2}.
\end{equation}
Taking nominal parameters of $r = 0.1 R_\star$, $x=1/2$, and
$\alpha^{\rm eff}_{\rm H\alpha}$=$10^{-13}$\,cm$^3$\,s$^{-1}$,
we find $n_{\rm H}\approx 5\times10^{11}$\,cm$^{-3}$.
Assuming a uniform density clump, this yields $M_{\rm H}\approx
8\times10^{16}$\,g.
A constraint on the magnetic field strength at the site of the clump
follows from the requirement that the magnetic pressure exceed the
thermal pressure, $B^2 / 8\pi > n_{\rm H} k T$.
Although we do not know the plasma temperature,
if it were significantly outside the range of
1,000 and 10,000\,K,
we would either full ionize the hydrogen, or not ionized enough of it.
We can therefore estimate
\begin{equation}
  B \gtrsim 3\,{\rm G}
  \left(
  \frac{n_{\rm H}}{5\times10^{11}\,{\rm cm}^{-3}}
  \frac{T}{5000\,{\rm K}}
  \right).
\end{equation}
Given typical  magnetic field strengths of low-mass stars, this bound
is easily met.
%TODO FIXME CITE KOCHUKHOV HERE

%%%%%%%%%%%%%%%%%%%%%%%%%
% Supplementary Figures %
%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\begin{figure}[!t]
  \centering
  \includegraphics[width=\textwidth]{figures/sf5.pdf}
  \caption{Time evolution of select regions in the Keck/HIRES
  spectra from 17 February 2025.
  The spectra have been smoothed with a Gaussian filter.
  The horizontal axis shows the relative velocity centered on each
  line's rest wavelength, normalized by the stellar equatorial
  velocity.  
  H$\alpha$ is the only Balmer line to show periodic variability of
  the form seen in Figure~\ref{fig:spec}.
  The He 5876\,\AA\ line shows a time-dynamic blueshift that 
  differs from the Balmer lines.
  Li 6708\,\AA\ shows no obvious absorption.
  }
  \label{fig:hirescuts}
\end{figure}

@@

Finally, the Li\,\textsc{I} 6708\,\AA\ doublet, which shows no obvious
absorption, as well as the broad K\,\textsc{I} 7699\,\AA\ resonance
line are both shown in Figure~\ref{fig:hirescuts}.  In the latter,
narrow telluric absorption features overlap the blue wing of the line.
Neither of these regions shows any notable variability.

@@

Circumstellar H$\alpha$ emission might be sourced either from
scattering of stellar H$\alpha$ photons, or from recombination.  We
neglect scattering because we see the amplitude of the circumstellar
emission to vary by a factor of $\approx$5, which is not mirrored in
the chromospheric line core.  The volume emissivity under case B
recombination (in which Lyman-series photons are trapped by high
optical depth, so that recombinations cascade through H$\alpha$ and
other Balmer lines) can be written as
\begin{equation}
  j_{\rm H\alpha} = n_{\rm e} n_{\rm p} \alpha^{\rm eff}_{\rm H\alpha}(T) h \nu_{\rm H\alpha},
\end{equation}
where $n_{\rm e}$ and $n_{\rm p}$ are the electron and proton densities,
and $\alpha^{\rm eff}_{\rm H\alpha}(T)$ is the effective recombination
coefficient, i.e.\ the number of recombinations per unit volume, per
unit time.
For hydrogen with temperatures between 1,000 and 10,000\,K with
densities $n_{\rm H}$$\sim$10$^{10}$\,cm$^{-3}$,
$\alpha^{\rm eff}_{\rm H\alpha}$ is typically on the order of $10^{-12}$ to
$10^{-13}$\,cm$^3$\,s$^{-1}$ \cite{Hummer1987}.
Neglecting the effects of atoms other than hydrogen, we can assume
an ionization fraction $x$, such that $n_{\rm e} = n_{\rm p} = x
n_{\rm H}$, for $n_{\rm H}$ the hydrogen number density.  Let $L_{\rm
H\alpha} = j_{\rm H\alpha} V$, for $V$ the volume of the emitting
hydrogen.
The luminosity of circumstellar hydrogen emission, $L_{\rm H\alpha}$,
is an observable: our SED fitting routine yields
$L_\star$$\approx$0.012\,$L_\odot$, which gives
$L_{\rm H\alpha},\star$$\approx$1.0$\times$$10^{28}$erg\,s$^{-1}$, and
the luminosity of the clump $L_{\rm H\alpha}$ is of order one tenth
that of the star.
Assuming an emitting sphere of radius $r$, we can write
\begin{equation}
  n_H = \frac{1}{x}
  \left( 
    \frac{L_{\rm H\alpha}}{ \alpha^{\rm eff}_{\rm H\alpha}(T) h \nu_{\rm H\alpha} V }
  \right)^{1/2}.
\end{equation}
Taking nominal parameters of $r = 0.1 R_\star$, $x=1/2$, and
$\alpha^{\rm eff}_{\rm H\alpha}$=$10^{-13}$\,cm$^3$\,s$^{-1}$,
we find $n_{\rm H}\approx 5\times10^{11}$\,cm$^{-3}$.
Assuming a uniform density clump, this yields $M_{\rm H}\approx
8\times10^{16}$\,g.
A constraint on the magnetic field strength at the site of the clump
follows from the requirement that the magnetic pressure exceed the
thermal pressure, $B^2 / 8\pi > n_{\rm H} k T$.
Although we do not know the plasma temperature,
if it were significantly outside the range of
1,000 and 10,000\,K,
we would either full ionize the hydrogen, or not ionized enough of it.
We can therefore estimate
\begin{equation}
  B \gtrsim 3\,{\rm G}
  \left(
  \frac{n_{\rm H}}{5\times10^{11}\,{\rm cm}^{-3}}
  \frac{T}{5000\,{\rm K}}
  \right).
\end{equation}
Given that the average surface magnetic field strengths of low-mass
stars have been measured to span hundreds to thousands of Gauss
\cite{Kochukhov2021}, this bound is easily met.
